Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-02-07 parallel-tractability-J-v9/conclusions.tex
proof read related work and conclusions

In this paper, we studied the problem of parallel tractability of
materialization on the datalog rewritable ontologies.
To identify the parallel tractability classes,
we proposed two \texttt{NC} algorithms, Algorithm~$\mathsf{A}_{bsc}^\psi$
and Algorithm~$\mathsf{A}_{opt}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding
parallel tractability classes such that materialization
on the datalog programs in these classes is in the NC complexity.
We further studied two specific ontology languages, DL-Lite and DHL (including one of its extension).
We showed that any ontology expressed in DL-Lite$_{core}$ or DL-Lite$_{\mathcal{R}}$ is tractable in parallel.
For DHL and DHL($\circ$), we proposed two restrictions such that materialization is tractable in parallel.

We analyzed the benchmark LUBM and the real-world dataset YAGO
and gave the cases where these ontologies belong to parallel tractability classes.
On the other hand, we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{prc}$, which can
also be restricted to an \texttt{NC} version.
We implemented a system based on Algorithm~$\mathsf{A}_{prc}$ and compared it to
the state-of-the-art reasoner RDFox.
The experimental results showed that our system is comparable to RDFox when handling
the LUBM and the YAGO ontologies.

In the future work, we will extend our work in two lines.
One line is a further study of the parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ones. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
tractable in parallel. Another line of the future work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other reasoning-based tasks,
like ontology classification, ontology debugging and query answering.

>>>>

In this paper, we studied the problem of parallel tractability of
materialization of datalog rewritable ontologies.
To identify classes that are tractable in parallel,
we proposed two \texttt{NC} algorithms, Algorithm~$\mathsf{A}_{bsc}^\psi$
and Algorithm~$\mathsf{A}_{opt}^\psi$, that perform materialization on
datalog rewritable ontology languages.
Based on these algorithms, we identified the corresponding classes of
datalog program for which materialization is tractable in
parallel (\texttt{PTD} classes), i.e., in the complexity class \texttt{NC}.
We further studied two specific ontology languages, DL-Lite and DHL (including an extension).
We showed that any ontology expressed in DL-Lite$_{\text{core}}$ or
DL-Lite$_{\mathcal{R}}$ is tractable in parallel.
For DHL and DHL($\circ$), we proposed two restrictions such that
materialization is tractable in parallel.

We verified the usefulness of our theoretical results in two ways. On the one hand,
we analyzed different kinds of datasets,
including well-known benchmarks, real-world ontologies and the famous YAGO dataset.
Our analysis showed that many real ontologies belong to
the class $\mathcal{D}_{\textit{\text{dhl}}}$ or $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$, for which
materialization is tractable in parallel. The
developers and users can also refer to $\mathcal{D}_{\textit{\text{dhl}}}$
and $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$
to create large-scale ontologies for which parallel tractability
is theoretically guaranteed. On the other hand,
we used an optimization strategy based on SWD paths
to give a practical algorithm variant Algorithm~$\mathsf{A}_{prc}$, which can
also be restricted to an \texttt{NC} version and can be implemented in practice.
We implemented a system based on Algorithm~$\mathsf{A}_{prc}$.
The experimental results showed that the optimizations proposed in this paper result
in a better performance on ontologies with parallel tractability compared
to the state-of-the-art reasoner RDFox.

In future work, we plan to extend our work in two lines.
One line is a further study of parallel tractability of
other expressive ontology languages, in addition to
datalog rewritable ones. Since different expressive OWL languages
have different syntaxes and higher reasoning complexities than \texttt{PTime}-complete, we need to
explore more restrictions that are practical and make materialization
tractable in parallel. Another line of work is to
study how to further apply the
results in practice.
One idea is to apply the technique of SWD paths to enhance other
reasoning-based tasks such as ontology classification, ontology debugging and query answering.

### 2018-02-07 parallel-tractability-J-v9/relatedWork.tex
proof read related work and conclusions

The parallel reasoner RDFox \cite{MotikNPHO14} is used to evaluate our implementation. RDFox is a state-of-the-art system that handles reasoning on datalog rewritable ontology languages. Algorithm~$\mathsf{A}_{bsc}$ proposed in Section~\ref{sec:ptclass} is similar to the main algorithm of RDFox (see \cite{MotikNPHO14}, Sections~3 and 4). The difference lies in that, a group of rule instantiations are handled by one processor (namely a thread) in RDFox, while in Algorithm~$\mathsf{A}_{bsc}$, each rule instantiation is assigned to a unique processor. Analogous to the method used in RDFox, ParallelDHL is also implemented by assigning a group of 
rule instantiations to a processor.
The experimental results show that ParallelDHL is comparable to RDFox when handling
the LUBM and the YAGO ontologies.

There is work that studies parallel reasoning in RDFS and OWL. The current methods mainly focus on optimizing reasoning algorithms from different aspects. The authors of \cite{GoodmanJMAAH11} propose a new kind of encoding method for RDF triples to achieve a high performance. This method can significantly yield a throughput improvement and optimize the parallel RDF reasoning and query answering. In the work of \cite{PetersSZ15}, the RETE algorithm is used to accelerate rule matching for RDFS reasoning. The authors of \cite{SubercazeGCL16} propose a more efficient storage technique and optimize the join operations in parallel reasoning. The issue of balance distribution of parallel tasks is also studied \cite{SomaP08,WeaverH09}. Two approaches are explored, i.e., \emph{rule partitioning} (allocating parallel tasks to different processors based on rules) and \emph{data partitioning} (allocating parallel tasks based on data). The evaluation results indicate that the efficiency of balance distribution varies with respect to different datasets. On the other hand, parallel reasoning is also implemented for OWL fragments, e.g., OWL~RL \cite{KolovskiWE10}, OWL~EL \cite{KazakovKS14}, OWL~QL \cite{LemboSS13}, and even highly expressive languages \cite{SteigmillerLG14,LiebigM07,SchlichtS08,WuH12}. In current work, several techniques are proposed to adapt parallel computation to OWL reasoning tasks. A kind of graph-based method is discussed in \cite{LemboSS13} to enhance OWL~QL classification. The authors of \cite{LiebigM07,SchlichtS08,WuH12} propose pruning techniques to optimize the Tableaux algorithm. The \emph{lock-free technique} is applied in the work \cite{KazakovKS14,SteigmillerLG14}.

Another line of optimizing parallel reasoning is to utilize high-performance computing platforms. For in-memory platforms, different supercomputers, like Cray XMT, Yahoo S4, have early been used in parallel RDF reasoning \cite{Hoeksema2011,GoodmanJMAAH11}. The authors of \cite{HeinoP12} report their work on RDFS reasoning based on massively parallel GPU hardware. The distributed parallel platforms, like MapReduce and Peer-to-Peer networks, are also used for RDFS reasoning. The representative systems are WebPIE \cite{UrbaniKMHB12}, Marvin \cite{oren2009marvin} and SAOR \cite{HoganHP09}. Different techniques are discussed in this work to tackle the special problems in distributed computing. However, to study the issue of parallel tractability on distributed platforms, we have to discuss more issues, e.g., \emph{network structures} and \emph{communications}. This is not the focus in this paper.

>>>>

The parallel reasoner RDFox \cite{MotikNPHO14} is used to evaluate our implementation. RDFox is a state-of-the-art system that handles reasoning on datalog rewritable ontology languages. Algorithm~$\mathsf{A}_{bsc}$ proposed in Section~\ref{sec:ptclass}, is similar to the main algorithm of RDFox (see \cite{MotikNPHO14}, Sections~3 and 4). The main difference is that a group of rule instantiations is handled by one processor (namely a thread) in RDFox, while in Algorithm~$\mathsf{A}_{bsc}$, each rule instantiation is assigned to a unique processor. Thus, in theory, the materialization of the datalog program in Example~\ref{exp:mg} is serial in RDFox {\color{red}The example has two rules. Why can they not be parallelized in RDFox?}. ItThe experiments also show that parallelism does not lead to a remarkable improvement for RDFox for the ontology series not in $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$. We use Algorithm~$\mathsf{A}_{opt}$ to show that the ontology in Example~\ref{exp:mg} is also tractable in parallel. Our implementation ParallelDHL, which is based on Algorithm~$\mathsf{A}_{prc}$, also shows a better performance than RDFox when handling the test ontologies that belong to $\mathcal{D}_{\textit{\text{dhl}}(\circ)}$.

There is work that studies parallel reasoning in RDFS and OWL. The current methods mainly focus on optimizing reasoning algorithms from different aspects. \citet{GoodmanJMAAH11} propose a new kind of encoding method for RDF triples to achieve a high performance. This method can significantly yield a throughput improvement and optimize parallel RDF reasoning and query answering. \citet{PetersSZ15} use the RETE algorithm to accelerate rule matching for RDFS reasoning. \citet{SubercazeGCL16} propose a more efficient storage technique and optimize the join operations in parallel reasoning. The issue of balance distribution of parallel tasks is also studied \cite{SomaP08,WeaverH09}. Two approaches are explored, i.e., \emph{rule partitioning} (allocating parallel tasks to different processors based on rules) and \emph{data partitioning} (allocating parallel tasks based on data). The evaluation results indicate that the efficiency of balance distribution varies with respect to different datasets. On the other hand, parallel reasoning is also implemented for OWL fragments, e.g., OWL~RL \cite{KolovskiWE10}, OWL~EL \cite{KazakovKS14}, OWL~QL \cite{LemboSS13}, and even highly expressive languages \cite{SteigmillerLG14,LiebigM07,SchlichtS08,WuH12}. In current work, several techniques are proposed to adapt parallel computation to OWL reasoning tasks. A kind of graph-based method is discussed by \citet{LemboSS13} to enhance OWL~QL classification.  Several pruning techniques are proposed to optimize the Tableaux algorithm \cite{LiebigM07,SchlichtS08,WuH12}. The \emph{lock-free technique} is applied by \citet{KazakovKS14} and \citet{SteigmillerLG14}.

Another line of optimizing parallel reasoning is to utilize high-performance computing platforms. For in-memory platforms, different supercomputers such as Cray XMT or Yahoo S4 have early on been used in parallel RDF reasoning \cite{Hoeksema2011,GoodmanJMAAH11}. \citet{HeinoP12} report on their work on RDFS reasoning based on massively parallel GPU hardware. Distributed parallel platforms such as MapReduce and Peer-to-Peer networks are also used for RDFS reasoning. The representative systems are WebPIE \cite{UrbaniKMHB12}, Marvin \cite{oren2009marvin} and SAOR \cite{HoganHP09}. Different techniques are discussed in this work to tackle the special problems in distributed computing. However, to study the issue of parallel tractability on distributed platforms,  more issues have to be discussed, e.g., \emph{network structures} and \emph{communications}. This is not the focus in this paper.
